---
title: "Fifa_Strikers"
author: "Shoaib"
date: "4/1/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plotly)
library(corrplot)
library(magick)
library(dplyr)
```

## R Markdown


```{r, echo=TRUE,warning=FALSE}
footballers <- feather::read_feather("/Users/shabdul/Desktop/Blog_Images/complete.feather")
nrow(footballers)
```
It contains 17,994 footballers from differnet nations and clubs

```{r, echo=TRUE}
strikers <- footballers[(footballers$prefers_st ==TRUE |
                           footballers$prefers_cf ==TRUE),
                        ]
nrow(strikers)
```
There are 3,409 strikers



## Age Distribution

Let's analyze age of strikers

```{r echo=TRUE,warning=FALSE}
plot_ly(x = strikers$age, opacity = 0.6, type = "histogram") %>%
  layout(barmode="overlay",
         title = "Age distribution of strikers",
         xaxis = list(title = "Age"),
         yaxis = list(title = "Number of Strikers")
  )
```




```{r}
summary(strikers$age)
```
We see the 3rd quartile for strikers age is 28

```{r,warning=FALSE}
# Let's look at strikers above or equal to 28 years of age
strikers_28orAbove <- footballers[(footballers$prefers_st ==TRUE |
                                     footballers$prefers_cf ==TRUE)
                                  & footballers$age >= 28 ,
                                  ]
nrow(strikers_28orAbove)
```

We see 975 strikers aged 28 or more


Let's look at correlation between various skill atttributes of these strikers

```{r,warning=FALSE,echo=TRUE}
# Let's try to look at the correlation between the skill atttributes of these strikers
correlation_dataframe <- as.data.frame(strikers_28orAbove[,sapply(strikers_28orAbove, function(x) is.numeric(x))])
drops <- c("ram","rcm","rdm","rcb","rwb","cam","cdm",
           "lwb","lam","lcm","ldm","lcb",
           "rw","rf","rs","rm","rb","st","lw","cf","cm","lm","lb","cb","ls","lf","gk",
           "gk_diving","gk_handling","gk_kicking","gk_positioning","gk_reflexes",
           "ID","eur_release_clause","potential",
           "def","marking","interceptions",
           "sliding_tackle","standing_tackle"
)
correlation_dataframe <- correlation_dataframe[ , !(names(correlation_dataframe) %in% drops)]
# Look at correlation plot
M <- cor(correlation_dataframe)
corrplot(M, method = "circle")
```

You notice some obvious trends here. Shoaib will call out a couple of  them here.... Pace, Stamina, Composure, Penalties and Freekicks seem interesting


```{r,warning=FALSE,echo=TRUE}
top_3_pace <- strikers_28orAbove %>%
  filter(rank(desc(pac))<=3)
img <- c(image_read(top_3_pace$flag[1]), image_read(top_3_pace$flag[2]), image_read(top_3_pace$flag[3]))
image_append(image_scale(img, "x100"))
```

```{r,warning=FALSE,echo=TRUE}
top_3_stamina<- strikers_28orAbove %>%
  filter(rank(desc(stamina))<=3)
img <- c(image_read(top_3_stamina$flag[1]), image_read(top_3_stamina$flag[2]), image_read(top_3_stamina$flag[3]))
image_append(image_scale(img, "x100"))
```

```{r,warning=FALSE,echo=TRUE}
plot_ly(strikers_28orAbove, x = ~age) %>%
  add_markers(y = ~stamina, text = ~name, color = I("red"), opacity=0.5, name = "Stamina", showlegend = TRUE) %>%
  add_markers(y = ~pac, text = ~name, color = I("blue"), opacity=0.5, name = "Pace", showlegend = TRUE) %>%
  add_lines(y = ~fitted(loess(pac ~ age)),
            line = list(color = '#1370D3',opacity=1),
            name = "Pace Trend", showlegend = TRUE) %>%
  add_lines(y = ~fitted(loess(stamina ~ age)),
            line = list(color = '#FF3F33',opacity=1),
            name = "Stamina Trend", showlegend = TRUE) %>%
  layout(xaxis = list(title = 'Age'),
         yaxis = list(title = 'Pace / Stamina'),
         title = "Impact of Age on Pace and Stamina",
         legend = list(x = 0.80, y = 0.90))
```

```{r,warning=FALSE,echo=TRUE}
top_2_penalties <- strikers_28orAbove %>%
  filter(rank(desc(penalties))<=3)
img <- c(image_read(top_2_penalties$flag[1]), image_read(top_2_penalties$flag[2]))
image_append(image_scale(img, "x100"))
```



```{r,warning=FALSE,echo=TRUE}
top_3_free_kick_accuracy<- strikers_28orAbove %>%
  filter(rank(desc(free_kick_accuracy))<=3)
img <- c(image_read(top_3_free_kick_accuracy$flag[1]), image_read(top_3_free_kick_accuracy$flag[2]), image_read(top_3_free_kick_accuracy$flag[3]))
image_append(image_scale(img, "x100"))
```


```{r,warning=FALSE,echo=TRUE}
plot_ly(strikers_28orAbove, x = ~age) %>%
  add_markers(y = ~penalties, text = ~name, color = I("red"), opacity=0.5, name = "Penalties", showlegend = TRUE) %>%
  add_markers(y = ~free_kick_accuracy, text = ~name, color = I("blue"), opacity=0.5, name = "Free kick accuracy", showlegend = TRUE) %>%
  add_lines(y = ~fitted(loess(penalties ~ age)),
            line = list(color = '#1370D3',opacity=1),
            name = "Penalties Trend", showlegend = TRUE) %>%
  add_lines(y = ~fitted(loess(free_kick_accuracy ~ age)),
            line = list(color = '#FF3F33',opacity=1),
            name = "Free kick_accuracy Trend", showlegend = TRUE) %>%
  layout(xaxis = list(title = 'Age'),
         yaxis = list(title = 'Penalties / Free_kick_accuracy'),
         title = "Impact of Age on Penalties and Free kick accuracy",
         legend = list(x = 0.80, y = 0.90))
```


```{r,warning=FALSE,echo=TRUE}
top_3_composure<- strikers_28orAbove %>%
  filter(rank(desc(composure))<=3)
img <- c(image_read(top_3_composure$flag[1]), image_read(top_3_composure$flag[2]), image_read(top_3_composure$flag[3]))
image_append(image_scale(img, "x100"))
```

```{r,warning=FALSE,echo=TRUE}
top_3_aggression<- strikers_28orAbove %>%
  filter(rank(desc(aggression))<=3)
img <- c(image_read(top_3_aggression$flag[1]), image_read(top_3_aggression$flag[2]), image_read(top_3_aggression$flag[3]))
image_append(image_scale(img, "x100"))
```



```{r,warning=FALSE,echo=TRUE}
plot_ly(strikers_28orAbove, x = ~age) %>%
  add_markers(y = ~composure, text = ~name, color = I("red"), opacity=0.5, name = "Composure", showlegend = TRUE) %>%
  add_markers(y = ~aggression, text = ~name, color = I("blue"), opacity=0.5, name = "Aggression", showlegend = TRUE) %>%
  add_lines(y = ~fitted(loess(composure ~ age)),
            line = list(color = '#1370D3',opacity=1),
            name = "Composure Trend", showlegend = TRUE) %>%
  add_lines(y = ~fitted(loess(aggression ~ age)),
            line = list(color = '#FF3F33',opacity=1),
            name = "Aggression Trend", showlegend = TRUE) %>%
  layout(xaxis = list(title = 'Age'),
         yaxis = list(title = 'Composure / Aggression'),
         title = "Impact of Age on Composure and Aggression",
         legend = list(x = 0.80, y = 0.90))
```

```{r,warning=FALSE,echo=TRUE}
strikers_above25_spread <- strikers_28orAbove %>% group_by(nationality)  %>% summarise(n = n()) %>% arrange(desc(n))
head(arrange(strikers_above25_spread, desc(n)), n=10)
```




```{r,warning=FALSE,echo=TRUE}
df <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/2014_world_gdp_with_codes.csv',stringsAsFactors = FALSE)
# adding 13 manual entries which are missing from the csv
df <- rbind(df[,c('COUNTRY','CODE')],c("England","GBR"),
            c("Bosnia Herzegovina","BIH"),
            c("DR Congo","COD"),
            c("Ivory Coast","CIV"),
            c("Cape Verde","CPV"),
            c("Republic of Ireland","IRL"),
            c("FYR Macedonia","MKD"),
            c("Korea Republic","KOR"),
            c("Congo","COG"),
            c("China PR","CHN"),
            c("Guinea Bissa","GNB"),
            c("Trinidad & Tobago","TTO"),
            c("Gambia","GMB"),
            c("Antigua & Barbuda","ATG")
)
location_mapping <- df %>% group_by(COUNTRY)  %>% summarise(n = CODE)
location_mapping <- data.frame(location_mapping)
strikers_above25_spread <- data.frame(strikers_above25_spread)
strikers_above25_spread$code <- location_mapping$n[match(strikers_above25_spread$nationality,location_mapping$COUNTRY)]
# adding light grey boundaries
l <- list(color = toRGB("grey"), width = 0.5)
# specifying map projection/options
g <- list(
  showframe = FALSE,
  showcoastlines = FALSE,
  projection = list(type = 'Mercator')
)
p <- plot_geo(strikers_above25_spread) %>%
  add_trace(
    z = ~n, color = ~n, colors = 'Blues',
    text = ~nationality, locations = ~code, marker = list(line = l)
  ) %>%
  colorbar(title = 'Population', tickprefix = '') %>%
  layout(
    title = 'Population Spread of Strikers above 25',
    geo = g
  )
p
```







